---
title: "p8105_hw6_av2718"
author: "Angelica Vina Albarracin"
date: "2022-12-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, message = FALSE)

library(dplyr)
library(tidyverse)
library(rvest)

```

# Problem 2

```{r}
#Load data raw data
homicide_raw = read_csv("data/homicide-data.csv") %>% 
  janitor::clean_names()

#Visualize first 8 rows
head(homicide_raw, 8) # view first 8 rows of raw data
```

This problem focuses on the `Homicide-data.csv` that The Washington Post gathered in 50 large U.S. cities and made available through GitHub. The raw data comprises `r dim(homicide_raw)`observations. The data included the geographical location of the homicide, whether an arrest was made, and, in most cases, basic demographic information about each victim. In total, the dataset contains 12 variables: `r ls(homicide_raw)`. Interestingly, the first and last names of the victims are included in the data, as well as the race of the victim, and exact location of the homicide (lat and log).

In the code chunk below, we create a `city_state` variable (e.g., "Baltimore, MD") and a binary variable `disposition` indicating whether the homicide was solved. We omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. 

```{r}
#Clean and arrange raw variables, create city_state variable, and make 'disposition' a binary variable

homicide_clean = homicide_raw %>% 
  mutate(
    city_state = str_c(city, ", " ,state), 
    disposition = if_else(disposition %in% "Closed by arrest", 1, 0),
    disposition = as.numeric(disposition),
     victim_age = as.numeric(victim_age)) %>% 
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"), #filter by city
         victim_race %in% c("White", "Black")) %>% #filter by race
  mutate(victim_race = fct_relevel(victim_race, "White")
         )%>% 
  select(!c(city, state))
    

#Visualize first 8 rows of new dataframe

head(homicide_clean, 8) # view first 8 rows of tidied data

```

For the city of Baltimore, in the code chunk below use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. To do this, we first filter our data set to select the Baltimore data:

```{r}
baltimore_data = homicide_clean %>% 
  filter(city_state == "Baltimore, MD") #new dataset with Baltimore data

#Fit logistic regression

glm1_baltimore = baltimore_data %>% 
  glm(disposition ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) 

glm1_results = glm1_baltimore %>% 
  broom::tidy(conf.int = T) %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value, conf.low, conf.high) 

print(glm1_results)

```

Now, we run glm for each of the cities in our dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.

```{r}

#nest data and run glm on each city

glm2_allcities = homicide_clean %>% 
  select(city_state, victim_race, victim_age, victim_sex, disposition) %>%
  nest(data = -city_state) %>% 
  mutate(models = map(data, ~glm(disposition ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())),
         results = map(models, ~broom::tidy(x = .x, conf.int = T))) %>% 
  select(-data, -models) %>% 
  unnest(results) %>% 
  filter(term %in% "victim_sexMale") %>% #filter by sex 
  mutate(OR = exp(estimate),
         conf.low = exp(conf.low),
         conf.high = exp(conf.high))
```

Lastly, we create a plot that shows the estimated ORs and CIs for each city:

```{r}

glm2_plot = glm2_allcities %>% 
  ggplot(aes(x = OR, y = reorder(city_state, OR)))+ #reorder cities according to proportion of homicides
  geom_point(color = 4)+
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high))+ # add error bars based on CIs
  labs(
    title = "Odds of Homicide Being Solved by Victim's Sex in US cities",
    x = "Odds Ratio",
    y = "US City" +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(axis.text.y = element_text(size = 6)) +
  theme(plot.title = element_text(hjust = 0)) +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_text(size = 7)) +
  theme(legend.title = element_text(size = 8)) +
  theme(legend.text = element_text(size = 6)) +
  theme(plot.title = element_text(hjust = 0.5))
  )

glm2_plot
```
In the plot, we see that in the majority of US cities, it is less likely to be a male victim in a solved homicide than being female, as the majority of cities have an odds ratio that is less than 1 for being a male victim. In Albuquerque, the odds of being a male victim in a solved homicide are significantly higher than that of a female victim. In contrast, in New York City, the odds ratio is the lowest in the US, and the odds of being a female victim in a solved homicide than a male victim is higher compared to other cities. It's important to note that there is a lot of variability in the data, and the CIs are large. Therefore, the data must be interpreted cautiously before drawing further conclusions. 

```

